---
title: "covid 19"
author: "Emann Malik"
date: "April 3, 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
  
```{r include=FALSE}
kRdata$Sex <- as.factor(kRdata$Sex)
kRdata$result <- as.factor(kRdata$result)
```
  
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex
  
$H_0$ Dying from COVID-19 is independent of sex.  
$H_A$ Dying from COVID-19 is not independent of sex.  
  
Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```
  
## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this
  
```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
The bar plots above show that, out of the people who survived COVID-19, there were more females (56%) than males (44%). Out of the people who did not survive COVID-19, there were more males (64%) than females (36%). The plots show that about twice as many males died from COVID-19 than females; there is an unequal amount of deaths between the two sexes. We are starting to see evidence that a person's sex could play a role in the chances of them dying from COVID-19.
  
## Numerical results
  
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
  
The tables above show quite a difference among survivals and deaths between males and females. Out of the females, more survived COVID-19 (56%) than those who did not (36%). Out of the males, more died from COVID-19 (64%) than those who survived (44%). More females lived than males and more males died than females. Once again, it appears that the virus is taking a larger toll on male lives.
  
## Inferential results 
  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
  
$H_0$ Dying from COVID-19 is independent of sex.     
$H_A$ Dying from COVID-19 is not independent of sex.    
  
The Chi-squared test above displays a p-value of 0.001 while Fisher's Exact Test displays a p-value of 0.0009. Both of these values indicate some sort of significance.
They, both representing the chance that we will get the same results just received if the null hypothesis is true, are small values.  
  
The odds ratio from Fisher's Exact Test, 2.3, is derived from the odds of a male dying (0.031) divided by the odds of a female dying (0.014). It means that males are 2.3 times more likely than females to die from COVID-19.
  
So, based on all these values, there is a significant difference between males and females; males are much more likely to die from COVID-19 than females. We can reject the null hypothesis in favor of the alternative: dying from COVID-19 is not independent of sex.
  
## Conclusions
  

  